---
title: "Debate_Important_Topics"
output: html_document
Author : "Vijay Pulijala"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Referred to the Important Topics in the debate 
In this short write-up, I'll look at the 4 important topics Jobs, Terrorism, immigration and health and how the the presidential candidates have used them during the first debate. 
* How often have the candidates used words similar to the 4 most important topics


```{r, message = F, warning = F, echo = F}

library(dplyr)
library(ggplot2)
library(scales)
library(tidytext)
debate <- read.csv("C:/Users/vpulij001c/Documents/Input/debate.csv", h =T, stringsAsFactors = F)

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```


```{r,echo = F,message = F, warning = F}
# English pronouns
jobs <- c("jobs","create")
terrorism<- c("isis", "ISIS","terrorism","terrorists","terrorist")
health <- c("education","health")
immigration <- c("immigration","immigrants","mexico","mexican","latina","latino","latinos","hispanic","hispanics")

# Bring them all together in a list
pronouns <- list("jobs" = jobs, "terrorism" = terrorism, "health" = health, "immigration" = immigration)


# Calculate word count & frequency
debate_words <- debate %>%
  filter(Speaker %in% c("Clinton", "Trump")) %>%
  unnest_tokens(word, Text) %>% # tidytext function
  group_by(word, Speaker) %>%
  summarise(count = n()) %>%
  group_by(Speaker) %>%
  mutate(frequency = count/sum(count)) %>%
  arrange(desc(frequency)) %>%
  ungroup()

# Create a pronoun variable using the list of pronouns
debate_words$pronoun <- ifelse(debate_words$word %in% pronouns$jobs, "jobs",
                               ifelse(debate_words$word %in% pronouns$terrorism, "terrorism",
                                      ifelse(debate_words$word %in% pronouns$health, "health",
                                          ifelse(debate_words$word %in% pronouns$immigration, "immigration",NA))))
```

The figure below shows the frequency of each important word pronoun as a percentage of each candidates' total words used. 

```{r, echo = F, message = F, warning = F}
jobs <- ggplot(debate_words %>% filter(pronoun == "jobs"), 
       aes(x = reorder(word, -frequency), y = frequency, fill = Speaker)) + 
  geom_bar(stat="identity", position="dodge", colour = "#333333") +
  scale_y_continuous(breaks = seq(0, .03, 0.005),
                     labels = percent) +
  labs(x = "Jobs", 
       y = "Percent of Candidates' Words", 
       title = "Reffered about Jobs") +
    scale_fill_manual(values = c("#4169E1", "#E91D0E")) +
    theme(plot.title = element_text(margin = margin(b = 10, unit = 'pt'),
                                  face = 'bold',
                                  colour = '#333333',
                                  size = rel(1.25),
                                  hjust = 0,
                                  vjust = 1))

terror <- ggplot(debate_words %>% filter(pronoun == "terrorism"), 
       aes(x = reorder(word, -frequency), y = frequency, fill = Speaker)) + 
  geom_bar(stat="identity", position="dodge", colour = "#333333") +
  
  scale_y_continuous(breaks = seq(0, .03, 0.005),
                     labels = percent) +
  
  labs(x = "Terrorism", 
       y = "Percent of Candidates' Words", 
       title = "Spoke on Terrorism") +
    scale_fill_manual(values = c("#4169E1", "#E91D0E")) +
    theme(plot.title = element_text(margin = margin(b = 10, unit = 'pt'),
                                  face = 'bold',
                                  colour = '#333333',
                                  size = rel(1.25),
                                  hjust = 0,
                                  vjust = 1))
health <- ggplot(debate_words %>% filter(pronoun == "health"), 
                 aes(x = reorder(word, -frequency), y = frequency, fill = Speaker)) + 
  geom_bar(stat="identity", position="dodge", colour = "#333333") +
  
  scale_y_continuous(breaks = seq(0, .03, 0.005),
                     labels = percent) +
  
  labs(x = "Health and Education", 
       y = "Percent of Candidates' Words", 
       title = "Health and Education") +
  scale_fill_manual(values = c("#4169E1", "#E91D0E")) +
  theme(plot.title = element_text(margin = margin(b = 10, unit = 'pt'),
                                  face = 'bold',
                                  colour = '#333333',
                                  size = rel(1.25),
                                  hjust = 0,
                                  vjust = 1))

immigration <- ggplot(debate_words %>% filter(pronoun == "immigration"), 
                 aes(x = reorder(word, -frequency), y = frequency, fill = Speaker)) + 
  geom_bar(stat="identity", position="dodge", colour = "#333333") +
  
  scale_y_continuous(breaks = seq(0, .03, 0.005),
                     labels = percent) +
  
  labs(x = "Immigration", 
       y = "Percent of Candidates' Words", 
       title = "Immigration") +
  scale_fill_manual(values = c("#4169E1", "#E91D0E")) +
  theme(plot.title = element_text(margin = margin(b = 10, unit = 'pt'),
                                  face = 'bold',
                                  colour = '#333333',
                                  size = rel(1.25),
                                  hjust = 0,
                                  vjust = 1))


multiplot(jobs, terror, health,immigration, cols = 2)
```

## Unique words used

The number of unique words used by the candidates increases monotonically over the course of the debate. But how do the trajectories in the cumulative sum of unique words differ for each candidate? Is Hillary firing off more flashy, low-frequency words compared to her opponent? 

```{r,echo = F, message = F, warning = F}
# Get cumulative sum of unique words over the transcript
debate_unique <- debate %>%
  filter(Speaker %in% c("Clinton", "Trump")) %>%
  unnest_tokens(word, Text) %>% 
  #select(-word_count) %>% # don't need word count
  group_by(Speaker) %>%
  mutate(cumulative_sum = cummax(as.numeric(factor(word, levels = unique(word))))) %>%
  arrange(Line, desc(cumulative_sum)) %>% distinct(Line, .keep_all = TRUE)    
```

The figure below shows us the total number of unique words used by each line in the transcript. 
It looks like Hillary may have a slight edge by the end of the debate, but it's not a wopping difference to me. Their unique word usage is pretty overlapped between lines 125 and 175. 

```{r, echo = F, message = F, warning = F}
# Plot cumulative sum of unique words for each candidate
ggplot(debate_unique, aes(x = Line, y = cumulative_sum, colour = Speaker)) + 
geom_line(lwd = 2, alpha = 0.8) +

scale_x_continuous(breaks = seq(0, nrow(debate), 25)) +
scale_y_continuous(breaks = seq(0, 1500, 250)) +

labs(x = "Time (Line in Transcript)", 
y = "Cumulative Sum of Unique Words Used", 
title = "Who is using a broader vocabulary during the debate?") +

scale_colour_manual(values = c("#4169E1", "#E91D0E")) +

theme(plot.title = element_text(margin = margin(b = 10, unit = 'pt'),
face = 'bold',
colour = '#333333',
size = rel(1.25),
hjust = 0,
vjust = 1))
```

